Journal and article coverage

### 2024-11-20 analysis/results.Rmd
First draft oa by publisher

## Combine to patchwork
layout <- "
AC
BB
"
wrap_plots(
  A = hoad_panel,
  B = wos_scp_uptake_panel,
  C = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(0.5, 1)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')

>>>>

## Combine to patchwork
layout <- "
AC
BB
"
wrap_plots(
  A = hoad_panel,
  B = wos_scp_uptake_panel,
  C = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(0.5, 1)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

### Scatterplot

```{r}
uptake_jn_five_jns <- uptake_df |>
  filter(earliest_year != "2018") |>
  group_by(issn_l, src) |>
  summarize(
    n_articles = sum(n_articles),
    n_oa_articles = sum(n_oa_articles),
    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
  )

uptake_scatter_df <- uptake_jn_five_jns |>
  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles)) |>
  select(issn_l, src, n_articles, n_oa_articles, author_position_oa = name, value) |>
  mutate(cat = tolower(paste(src, author_position_oa, sep = "_")))
```

```{r}
plot_uptake_jn_five_jns <- uptake_jn_five_jns |>
  mutate(oa_ta_first = ta_oa_first_author_articles,
         oa_ta_corr =  ta_oa_corresponding_author_articles) |> 
  distinct(issn_l, src, oa_ta_first, oa_ta_corr) |> 
  pivot_wider(names_from = src,
              values_from = c("oa_ta_first", "oa_ta_corr")) |> 
  janitor::clean_names() |>
  pivot_longer(
    cols = c(oa_ta_first_scopus, oa_ta_first_web_of_science, oa_ta_corr_scopus, oa_ta_corr_web_of_science)
  ) |>
  distinct(issn_l, oa_ta_first_hoad, name, value)

ggplot(plot_uptake_jn_five_jns, aes(oa_ta_first_hoad, value, group = name))  +
  geom_point(alpha = 0.05) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
  coord_fixed() +
  facet_grid(~name)
```

```{r}
uptake_publisher <- uptake_df |>
  inner_join(esac, by = "issn_l") |>
  filter(earliest_year != "2018")

uptake_publisher_by_year <- uptake_publisher |>
  mutate(earliest_year = as.character(earliest_year)) |>
  group_by(earliest_year, src, publisher) |>
  summarize(
    n_articles = sum(n_articles),
    n_oa_articles = sum(n_oa_articles),
    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
  )

uptake_by_year_publisher_df <- uptake_publisher_by_year |>
  # Remove HOAD because we only have first author data available
  filter(src != "HOAD", earliest_year != "2018") |>
  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))

hoad_publisher_panel <- uptake_publisher_by_year |>
  filter(src == "HOAD") |>
  ggplot(aes(x = earliest_year)) +
  geom_bar(
    aes(y = n_oa_articles / n_articles, fill = "All"),
    stat = "identity",
    color = "#000000"
  ) +
  geom_bar(
    aes(y = ta_oa_first_author_articles / n_articles, fill = "TA OA"),
    stat = "identity",
    color = "#000000"
  ) +
  geom_text(
    aes(
      label = paste0(
        round(ta_oa_first_author_articles / n_oa_articles * 100, 1),
        "%\n(",
        format(ta_oa_first_author_articles, big.mark = ","),
        ")"
      ),
      y = ta_oa_first_author_articles / n_articles),
    position = position_dodge(width = 1),
    vjust = -0.1,
    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
    size = 0.8 * my_base_size / .pt
  ) + 
  facet_grid(~fct_rev(publisher)) +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#56B4E9"),
    name = "",
    guide = guide_legend(reverse = TRUE)
  ) +
   scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .4),
    breaks = c(0, .1, .20, .3, .4)
  ) +
  theme_minimal(base_size = my_base_size) +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left"
  )

## label test

first_au_publisher_panel <- uptake_by_year_publisher_df |>
  distinct(earliest_year, publisher, n_oa_articles, n_articles, src, name, value) |>
  # Modify factor levels to control order
  mutate(
    src = factor(src, levels = c("Web of Science", "Scopus")),
    name = factor(name, 
                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
  ) |>
  filter(name == "ta_oa_first_author_articles") |>
  ggplot() +
   geom_bar(
    aes(
      x = earliest_year,
      y = n_oa_articles / n_articles,
      fill = "All"
    ),
    stat = "identity",
    color = "#000000a0",
    size = .2
  ) +
  geom_bar(
    aes(
      x = earliest_year,
      y = value / n_articles,
      fill = "TA OA"
    ),
    stat = "identity",
    color = "#000000a0",
    size = .2
  ) +
   geom_text(
    aes(
      label = paste0(
        round(value / n_oa_articles * 100, 1),
        "%\n(",
        format(value, big.mark = ","),
        ")"
      ),
      x = earliest_year,
      y = value / n_articles),
    position = position_dodge(width = 1),
    vjust = -0.2,
    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
    size = 0.7 * my_base_size / .pt
  ) +
  facet_grid(src ~ fct_rev(publisher)) +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#56B4E9"),
    name = "",
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .4),
    breaks = c(0, .1, .20, .3, .4)
  ) +
  theme_minimal(base_size = my_base_size) +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left")

cor_au_publisher_panel <- uptake_by_year_publisher_df |>
  distinct(earliest_year, publisher, n_oa_articles, n_articles, src, name, value) |>
  # Modify factor levels to control order
  mutate(
    src = factor(src, levels = c("Web of Science", "Scopus")),
    name = factor(name, 
                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
  ) |>
  filter(name == "ta_oa_corresponding_author_articles") |>
  ggplot() +
  geom_bar(
    aes(
      x = earliest_year,
      y = n_oa_articles / n_articles,
      fill = "All"
    ),
    stat = "identity",
    color = "#000000",
    size = .2
  ) +
  geom_bar(
    aes(
      x = earliest_year,
      y = value / n_articles,
      fill = "TA OA"
    ),
    stat = "identity",
    color = "#000000",
    size = .2
  ) +
   geom_text(
    aes(
      label = paste0(
        round(value / n_oa_articles * 100, 1),
        "%\n(",
        format(value, big.mark = ","),
        ")"
      ),
      x = earliest_year,
      y = n_oa_articles / n_articles),
    position = position_dodge(width = 1),
    vjust = -0.1,
    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
    size = 0.8 * my_base_size / .pt
  ) +
  facet_grid(src ~ fct_rev(publisher)) +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#56B4E9"),
    name = "",
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .4),
    breaks = c(0, .1, .20, .3, .4)
  ) +
  theme_minimal(base_size = my_base_size) +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left")
```

```{r, fig.asp=0.7}
## Combine to patchwork
layout <- "
A
B
C
"
wrap_plots(
  A = hoad_publisher_panel,
  B = first_au_publisher_panel,
  C = cor_au_publisher_panel,
  D = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(1, 2, 2)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

- Elsevier gap suggests different oa information, let's make a scatterplot and contrast it hoad info using a shared corpus. due to publisher-specifc license

ta by country
